---
title: "SSL Validation"
output: github_document
---

## Overview

Phenotyping algorithms based on Electronic Health Records (EHR) aim to identify a patient’s disease status using the information in the health record.

Algorithm evaluation in EMR data is often based on a small number of gold-standard labeled data. Two main issues are the difficulty in selecting the appropriate cut and the high variance in the accuracy parameter estimates.

This package provides a semi-supervised learning approach that incorporates unlabeled data into the estimation of the receiver operating characteristic (ROC) curve parameters to address these issues. 

The data consists of :

- a small validation set with algorithm score S and label Y, which takes values 0 or 1
	
- a large unlabeled set containing only the algorithm score S.

The main function `roc.semi.superv` takes as arguments S and Y and provides a semi-supervised estimation of the ROC parameters. 

The function `roc.superv` takes as arguments S and Y from the small validation set only and provides a classic supervised method to estimate the ROC parameters.

## Installation

Install development version from GitHub.

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("celehs/SSL.validation")
```

Load the package into R.

```{r}
library(SSL.validation)
```

## Simulated Example

```{r}
set.seed(1234)
dat <- read.csv("https://raw.githubusercontent.com/celehs/SSL.validation/master/data-raw/data.csv")
```

```{r}
p.0 <- mean(dat$Y , na.rm = TRUE)
id.v <- which(is.na(dat$Y) != 1)
dat.v <- dat[id.v, ] # Labeled Data 
```

### Semi-Supervised Learning (SSL)

```{r}
system.time(res.ssl <- roc.semi.superv(dat$S,dat$Y))
auc.ssl <- res.ssl$auc
roc.ssl <- res.ssl$roc
auc.ssl
tail(roc.ssl)
```

### Supervised Learning (SL)

```{r}
system.time(res.sl <- roc.superv(dat.v$S,dat.v$Y))
auc.sl <- res.sl$auc
roc.sl <- res.sl$roc
colnames(roc.sl) <- colnames(roc.ssl)
auc.sl
tail(roc.sl)
```
